Efficient and Scalable Bipartite Matching with Fast Beta Linkage (fabl)

### 2023-07-20 ba_resubmit/kundinger_final.tex
putting jerry's proposed changes in red text.

\begin{abstract} 
Within the field of record linkage, Bayesian methods have the crucial advantage of quantifying uncertainty from imperfect linkages. However, current implementations of Bayesian Fellegi-Sunter models are computationally intensive, making them challenging to use on larger-scale record linkage tasks. To address these computational considerations, we propose fast beta linkage (\texttt{fabl}), an extension to the Beta Record Linkage (\texttt{BRL}) method of \cite{sadinle_bayesian_2017}. Specifically, we use independent prior distributions over the matching space, \textcolor{teal}{which allows for parallel updates to the Gibbs sampler, contrasting the sequential updates of \texttt{BRL}.}
\textcolor{teal}{We ensure our method is bipartite using decision theoretic post-processing techniques from the existing literature.}
\textcolor{teal}{Next, we use hashing techniques that reduce computational overhead. Furthermore, we complete pairwise record comparisons over large data files through parallel computing and reduce memory costs through a new technique called storage efficient indexing. 
}
%This allows
%us to use hashing techniques that reduce computational overhead. This also allows us to complete pairwise record comparisons over large data files through parallel computing and to reduce memory costs through a new technique called storage efficient indexing. 
Through simulations and two case studies, we show that \texttt{fabl} can have markedly increased speed with minimal loss of accuracy when compared to \texttt{BRL}.
\end{abstract}

%% ** Keywords **
\begin{keyword}
\kwd{bipartite record linkage}
\kwd{entity resolution}
\kwd{data cleaning}
\kwd{parallelization}
\kwd{hashing}
\end{keyword}

@@

In this article, we propose fast beta linkage (\texttt{fabl}), which extends the \texttt{BRL} model for increased efficiency and scalability. {\color{red}{We use independent priors for the matching status of each record, inducing a dependency that is stronger than the record pair independence in the original \cite{fellegi_theory_1969} but weaker that the one-to-one requirement of \cite{sadinle_bayesian_2017}.}} \textcolor{teal}{This independence assumption allows for a scalable update of the Gibbs sampler, contrasting sequential updates of \cite{sadinle_bayesian_2017}. We employ the decision theoretic technique of \cite{sadinle_bayesian_2017} in order to ensure our method after record linkage is bipartite.}
This \textcolor{teal}{discovery} allows us to (1) employ hashing techniques that speed up calculations and reduce computational costs, (2) compute the pairwise record comparisons over large data files via parallel computing, and (3) reduce memory costs through what we call storage efficient indexing. These contributions allow \texttt{fabl} to perform record linkage on much larger data files than previous Bayesian Fellegi-Sunter models at significantly increased speed with similar levels of accuracy. In particular, computation time under \texttt{BRL} grows quadratically, with the size of each data file, while computation time under \texttt{fabl} grows linearly, only with the size of the smaller data file.

@@

\cite{sadinle_bayesian_2017} presents a Bayesian version of the model in \eqref{eqn:fs_model} through \eqref{eqn:lambda_dist}.  He uses uniform Dirichlet prior distributions for each $\bm{m}_f$ and $\bm{u}_f$.  As the prior distribution for  $\bm{Z}$, he proposes the ``beta distribution for bipartite matching."  This assigns a Bernoulli distribution for the indicator that a record in $B$ has a match in $A$, that is, $I(Z_j \leq n_A) \sim \text{Bernoulli}(\pi)$.  Additionally, $\pi \sim \text{Beta}(\alpha_{\pi}, \beta_{\pi})$, \textcolor{teal}{where  $\alpha_{\pi}, \beta_{\pi}$ are known hyper-parameters.}

>>>>

\begin{abstract} 
Within the field of record linkage, Bayesian methods have the crucial advantage of quantifying uncertainty from imperfect linkages. However, current implementations of Bayesian Fellegi-Sunter models are computationally intensive, making them challenging to use on larger-scale record linkage tasks. To address these computational considerations, we propose fast beta linkage (\texttt{fabl}), an extension to the Beta Record Linkage (\texttt{BRL}) method of \cite{sadinle_bayesian_2017}. Specifically, we use independent prior distributions over the matching space, \textcolor{teal}{which allows for parallel updates to the Gibbs sampler, \textcolor{red}{unlike} the sequential updates of \texttt{BRL}.}
\textcolor{teal}{We ensure \textcolor{red}{\text{fabl} results in a bipartite matching} using decision theoretic post-processing techniques from the existing literature.}
\textcolor{red}{To reduce computational overhead and memory costs, we use hashing and a new technique that we call storage efficient indexing. }
%\textcolor{teal}{Next, we use hashing techniques that reduce computational overhead. Furthermore, we complete pairwise record comparisons over large data files through parallel computing and reduce memory costs through a new technique called storage efficient indexing. 
%}
%This allows
%us to use hashing techniques that reduce computational overhead. This also allows us to complete pairwise record comparisons over large data files through parallel computing and to reduce memory costs through a new technique called storage efficient indexing. 
Through simulations and two case studies, we show that \texttt{fabl} can have markedly increased speed with minimal loss of accuracy when compared to \texttt{BRL}.
\end{abstract}

%% ** Keywords **
\begin{keyword}
\kwd{record linkage}
\kwd{entity resolution}
\kwd{data cleaning}
\kwd{parallelization}
\kwd{hashing}
\end{keyword}

@@

In this article, we propose fast beta linkage (\texttt{fabl}), which extends the \texttt{BRL} model for increased efficiency and scalability. 
\textcolor{red}{We use independent prior distributions for the matching status of each record, an idea suggested in \cite{wortman2019}.}
%% Jerry thought the wording below was too vague and made a suggestion that is more direct and clear. 
%{\color{red}{We use independent priors for the matching status of each record, inducing a dependency that is stronger than the record pair independence in the original \cite{fellegi_theory_1969} but weaker that the one-to-one requirement of \cite{sadinle_bayesian_2017}.}} 
%
\textcolor{teal}{This independence assumption allows for a scalable update of the Gibbs sampler, \textcolor{red}{unlike the} sequential updates of \cite{sadinle_bayesian_2017}. We employ the decision theoretic technique of \cite{sadinle_bayesian_2017} in order to ensure our method after record linkage is bipartite.}
This \textcolor{red}{approach} allows us to (1) employ hashing techniques that speed up calculations and reduce computational costs, (2) compute the pairwise record comparisons over large data files via parallel computing, and (3) reduce memory costs through what we call storage efficient indexing. These contributions allow \texttt{fabl} to perform record linkage on much larger data files than previous Bayesian Fellegi-Sunter models at significantly increased speed with similar levels of accuracy. In particular, computation time under \texttt{BRL} grows quadratically, with the size of each data file, while computation time under \texttt{fabl} grows linearly, only with the size of the smaller data file.

@@

\cite{sadinle_bayesian_2017} presents a Bayesian version of the model in \eqref{eqn:fs_model} through \eqref{eqn:lambda_dist}.  He uses uniform Dirichlet prior distributions for each $\bm{m}_f$ and $\bm{u}_f$.  \textcolor{red}{However, he replaces (\ref{eqn:lambda_dist}) with a prior distribution for $\mathbf{Z}$ that he calls the ``beta distributions for bipartite matching."}
%
%As the prior distribution for  $\bm{Z}$, he proposes the ``beta distribution for bipartite matching."  
This assigns a Bernoulli distribution for the indicator that a record in $B$ has a match in $A$, that is, $I(Z_j \leq n_A) \sim \text{Bernoulli}(\pi)$.  Additionally, $\pi \sim \text{Beta}(\alpha_{\pi}, \beta_{\pi})$, \textcolor{teal}{where  $\alpha_{\pi}, \beta_{\pi}$ are known hyper-parameters.}
